Reproducible Research: Peer Assessment 1
========================================================

```{r download, echo=FALSE}
#Download & Unzip Activity monitoring Dataset
setwd("C:/Users/js1/Dropbox/Repository/RepData_PeerAssessment1")
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(fileUrl, destfile = "Dataset.zip")
#unzip("Dataset.zip")
```

* Loading and preprocessing the data
```{r}
activity = read.csv("activity.csv", sep=",", na.strings = "NA")
activity$date <- as.Date(activity$date)
```

* Summary of the Dataset
```{r}
summary(activity)
```

* Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
q<-qplot(date, weight=activity$steps, data=activity, geom="histogram")
suppressMessages(print(q)) #Supressing ggplot messages
```

* Calculate and report the mean and median total number of steps taken per day
```{r}
mean <- mean(tapply(activity$steps, activity$date, sum, na.rm = TRUE))
median <- median(tapply(activity$steps, activity$date, sum, na.rm = TRUE))
```

The Mean is: **`r mean`** and the Median: **`r median`**

* Time Series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r average_steps, fig.width=10, fig.height=5}
average_steps<-data.frame(cbind(activity$interval,tapply(activity$steps, activity$interval, mean, na.rm = TRUE)))
colnames(average_steps) <- c("interval","steps")

q<-ggplot(data=average_steps,aes(x=interval,y=steps, color=steps)) +
    geom_line() 

print(q)
```

* The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps
```{r}
maxtime<-average_steps[which.max(average_steps$steps),"interval"]
strftime( as.POSIXct(Sys.Date()) + as.difftime(round(maxtime/100), units="hours")+ as.difftime(maxtime%%100, units="mins"), "%r",tz="UTC")
```

* Imputing missing values
```{r}
missing <- sum(is.na(activity))
```

Number of missing values is `r missing`

* Filling in all of the missing values in the dataset by using mean for that 5-minute interval
```{r}
#Copy of the original DataSet
fixed <- activity

#Remove duplicates 
average_steps <- average_steps[!duplicated(average_steps),]

for (i in 1:nrow(fixed)) {
    if (is.na(fixed[i, "steps"])){
        x <- fixed[i, "interval"]
        temp <- average_steps[average_steps$interval == x,]
        fixed[i,"steps"] <- temp$steps
        }
}
```

* Histogram of the total number of steps taken each day 
```{r total_steps, fig.width=10, fig.height=5}
q<-qplot(date, weight=fixed$steps, data=fixed, geom="histogram")
suppressMessages(print(q)) #Supressing ggplot messages
```

* Calculate and report the mean and median total number of steps taken per day. 
```{r}
mean_fixed <- mean(tapply(fixed$steps, fixed$date, sum, na.rm = TRUE))
median_fixed <- median(tapply(fixed$steps, fixed$date, sum, na.rm = TRUE))
```        

* Do these values differ from the estimates from the first part of the assignment?: 
```r (mean!=mean_fixed && median!=median_fixed)```

* What is the impact of imputing missing data on the estimates of the total daily number of steps?

New Mean: **`r mean_fixed`**        First mean: **`r mean`**
New Median: **`r median_fixed`**    First Median: **`r median`**

* Create New factor variable in the dataset with two levels - "weekday" and "weekend" 
```{r}
#install.packages("timeDate")
library(timeDate)

for (i in 1:nrow(fixed)) {
    if (isWeekend(fixed[i, "date"])){
        #print(fixed[i, "date"])
        fixed[i, "wk"] <- "weekend"
    } else {
        fixed[i, "wk"] <- "weekday"
    }
}

fixed$wk <- as.factor(fixed$wk)
```

* Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r , fig.width=10, fig.height=5}
q<-ggplot(data=fixed,aes(x=interval,y=steps)) + facet_grid(wk ~ .) + geom_line()
print(q)
```


